---
title: "Network R for Data Mining Book"
author: "Johnny Zhang"
date: "December 16, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(intergraph) ## asIgraph, asNetwor library(sna)
library(network)
library(sna)

library(tidyverse)
library(gplots) ## heatmap
```

## Data

We first create a network object with the covariates.

```{r}
## nfriends, imputed using mean for one student
nonnetwork <- read.csv("t1_nonnetwork.csv")

wechat <- read.csv("t1_wechat.csv")
wechat <- symmetrize(wechat, rule="weak")

wechat.network <- network(wechat, directed=F)

## adding the covariates
depress <- nonnetwork %>% select(depress1:depress7) %>% rowSums(na.rm = T)
lonely <- nonnetwork %>% select(sad1:sad10) %>% rowSums(na.rm = T)
happy <- nonnetwork %>% select(happy1:happy4) %>% rowSums(na.rm = T)
set.vertex.attribute(wechat.network, "depress", depress)
set.vertex.attribute(wechat.network, "lonely", lonely)
set.vertex.attribute(wechat.network, "happy", happy)

set.vertex.attribute(wechat.network, "gpa", nonnetwork$gpa)
set.vertex.attribute(wechat.network, "age", nonnetwork$age)
set.vertex.attribute(wechat.network, "gender", as.character(nonnetwork$gender))
set.vertex.attribute(wechat.network, "smoke", as.character(nonnetwork$smoke))
set.vertex.attribute(wechat.network, "drink", nonnetwork$drink)
set.vertex.attribute(wechat.network, "nfriend", nonnetwork$wechat)
bmi <- nonnetwork$weight/(nonnetwork$height/100)^2
set.vertex.attribute(wechat.network, "bmi", bmi)

summary(wechat.network, print.adj = F)
```

## Data visualization

### heat map

```{r}
wechat.mat <- as.matrix(wechat)
pdf("heatmap.pdf", width=7, height=7)
par(mar=c(0,1,1,0))
heatmap.2(wechat.mat, dendrogram='none', Rowv=F, Colv=F,trace='none', key=F, lhei = c(0.05, 0.95), lwid=c(.05, .95), col=colorpanel(n=2,low="grey", high="black"))
dev.off()

pdf("heatmap.ordered.pdf", width=7, height=7)
par(mar=c(0,1,1,0))
heatmap.2(wechat.mat, dendrogram='none', Rowv=TRUE, Colv="Rowv", trace='none', key=F, lhei = c(0.05, 0.95), lwid=c(.05, .95), col=colorpanel(n=2,low="grey", high="black"))
dev.off()
```


### Network plot

```{r}
## plot the whole network
pdf("whole.network.pdf", width=7, height=7)
par(mar=c(0,0,0,1))
set.seed(3)
layout.coord <- gplot(wechat.network, gmode="graph", usearrows=F, vertex.cex=1.5, vertex.col='black', edge.col='grey75', mode="fruchtermanreingold")
dev.off()

pdf("whole.network.mds.pdf", width=7, height=7)
par(mar=c(0,0,0,1))
layout.coord <- gplot(wechat.network, gmode="graph", usearrows=F, vertex.cex=1.5, vertex.col='blue', edge.col='grey75', mode="mds")
dev.off()

#save(layout.coord, file="all.coord.RData")

## plot the groups
## according to gender
pdf("whole.network.gender.pdf", width=7, height=7)
par(mar=c(0,0,0,1))
gender.col <- ifelse(nonnetwork$gender==1, "blue", "red") ## 1 female; 0 male
gplot(wechat.network, gmode="graph", usearrows=F, vertex.cex=1.5, vertex.col=gender.col, edge.col='grey75', coord = layout.coord)
legend("topright", c("Female", "Male"), col=c("blue", "red"), pch=c(16,16), cex = c(1.5, 1.5))
dev.off()

## according to class
pdf("whole.network.class.pdf", width=7, height=7)
par(mar=c(0,0,0,1))
gplot(wechat.network, gmode="graph", usearrows=F, vertex.cex=1.5, vertex.col=nonnetwork$class, edge.col='grey75', coord = layout.coord)
legend("topright", c("A", "B", "C", "D", "E", "F"), col=1:6, pch=rep(16,6), cex = rep(1.5, 6))
dev.off()

## plot a subset of data
# Take out a subset of the network - the first 15 students
wechat.sub.net <- wechat.network
wechat.sub.net <- delete.vertices(wechat.sub.net, 16:165)
summary(wechat.sub.net)

pdf("sub.network.pdf", width=7, height=7)
par(mar=c(1,1,1,1))
set.seed(2)
layout.coord <- gplot(wechat.sub.net, gmode="graph", usearrows=F, vertex.cex=1.8, vertex.col='black', edge.col='grey75', mode="fruchtermanreingold", displaylabels=T, label.pos=5, label.col = "white")
dev.off()

pdf("network-4-1.pdf", width=7, height=7)
par(mar=c(1,1,1,1))
set.seed(2)
layout.coord <- gplot(wechat.sub.net, gmode="graph", usearrows=F, vertex.cex=1.8, vertex.col='black', edge.col='grey75', mode="fruchtermanreingold", displaylabels=T, label.pos=5, label.col = "white")
dev.off()

pdf("network-4-2.pdf", width=7, height=7)
par(mar=c(1,1,1,1))
layout.coord <- gplot(wechat.sub.net, gmode="graph", usearrows=F, vertex.cex=1.8, vertex.col='black', edge.col='grey75', mode="circle", displaylabels=T, label.pos=5, label.col = "white")
dev.off()

pdf("network-4-3.pdf", width=7, height=7)
par(mar=c(1,1,1,1))
layout.coord <- gplot(wechat.sub.net, gmode="graph", usearrows=F, vertex.cex=1.8, vertex.col='black', edge.col='grey75', mode="mds", displaylabels=T, label.pos=5, label.col = "white")
dev.off()

pdf("network-4-4.pdf", width=7, height=7)
par(mar=c(1,1,1,1))
layout.coord <- gplot(wechat.sub.net, gmode="graph", usearrows=F, vertex.cex=1.8, vertex.col='black', edge.col='grey75', mode="random", displaylabels=T, label.pos=5, label.col = "white")
dev.off()

## plot the groups
## according to gender
pdf("sub.network.gender.pdf", width=7, height=7)
par(mar=c(1,1,1,1))
gender.col <- ifelse(nonnetwork$gender[1:15]==1, "blue", "red") ## 1 female; 0 male
gplot(wechat.sub.net, gmode="graph", usearrows=F, vertex.cex=1.8, vertex.col=gender.col, edge.col='grey75', coord = layout.coord, displaylabels=T, label.pos=5, label.col = "white")
legend("topright", c("Female", "Male"), col=c("blue", "red"), pch=c(16,16), cex = c(1.5, 1.5))
dev.off()
```

## Network Statistics
### Whole network
```{r}
## density
gden(wechat.network)

## transitivity
gtrans(wechat.network)
```

### node statistics

```{r}
nodedegree <- degree(wechat.network, cmode="indegree")
nodedegree
which.max(nodedegree)
which.min(nodedegree)
mean(nodedegree)

pdf("node-degree.pdf", width=7, height=7)
hist(nodedegree, xlab="Degree", main="")
dev.off()
```

```{r}
incloseness = closeness(wechat.network, cmode = "suminvundir")
incloseness

index <- 1:165

index[which.max(incloseness)]
index[which.min(incloseness)]

pdf("node-closeness.pdf", width=7, height=7)
hist(incloseness, xlab="Closeness", main="")
dev.off()
```

```{r}
node.betweenness <- betweenness(wechat.network, cmode = "undirected")
node.betweenness

index <- 1:165

index[which.max(node.betweenness)]
index[which.min(node.betweenness)]

pdf("node-betweenness.pdf", width=7, height=7)
hist(node.betweenness, xlab="Betweenness", main="")
dev.off()
```

#### page rank

```{r}
node.page.rank <- igraph::page.rank(asIgraph(wechat.network))$vector

node.page.rank

index[which.max(node.page.rank)]
index[which.min(node.page.rank)]

pdf("node-page.rank.pdf", width=7, height=7)
hist(node.page.rank, xlab="PageRank", main="")
dev.off()
```
### Dyadic statistics

the shortest paths from or to the vertices in the network.

```{r}
shortestpath <- igraph::distances(asIgraph(wechat.network))

#shortestpath
max(shortestpath)

table(shortestpath)/(165*164)

pdf("shortestpath.pdf", width=7, height=7)
par(mar=c(0,1,1,0))
heatmap.2(shortestpath, dendrogram='none', Rowv=F, Colv=F,trace='none', key=F, lhei = c(0.05, 0.95), lwid=c(.05, .95), col=colorpanel(n=5,low="white", high="black"))
dev.off()

pdf("shortestpath.pdf", width=7, height=7)
par(mar=c(0,1,1,0))
heatmap.2(shortestpath, dendrogram='none', Rowv=F, Colv=F,trace='none', key=F, lhei = c(0.05, 0.95), lwid=c(.05, .95), col=1:5)
#text(5:8, 10, labels=1:4, col=2:5)
#legend('topright', legend=0:4, col=1:5)
dev.off()

pdf("shortestpath.pdf", width=7, height=7)
par(mar=c(0,1,1,0))
heatmap.2(shortestpath, dendrogram='none', Rowv=F, Colv=F,trace='none', key=T, col=1:5, density.info="none", key.title="", key.xlab = "", keysize=0.8, lhei = c(0.2, 0.8), lwid=c(.2, .8))
#text(5:8, 10, labels=1:4, col=2:5)
#legend('topright', legend=0:4, col=1:5)
dev.off()

pdf("shortestpath.pdf", width=7, height=7)
par(mar=c(0,1,1,0))
heatmap.2(shortestpath, dendrogram='none', Rowv=F, Colv=F,trace='none', key=T, col=1:5, density.info="none", key.title="", key.xlab = "", keysize=1)
#text(5:8, 10, labels=1:4, col=2:5)
#legend('topright', legend=0:4, col=1:5)
dev.off()
```

## Basic network analysis

### t-test

```{r}
dset1 <- as.data.frame(cbind(gender = nonnetwork$gender,
                    gpa = nonnetwork$gpa,
                    bmi = bmi,
                    wechat = log(nonnetwork$wechat+1),
                    smoke = nonnetwork$smoke,
                    drink = nonnetwork$drink,
                    depress = depress,
                    happy = happy,
                    lonely = lonely,
                    degree = log(nodedegree),
                    closseness = log(incloseness),
                    betweenness = log(node.betweenness),
                    pagerank = log(node.page.rank),
                    major = nonnetwork$class,
                    age = nonnetwork$age
                    ))
```

```{r}
## gender difference in degree
t.test(degree ~ gender, data=dset1)

pdf("degree-boxplot.pdf", width=7, height=7)
boxplot(degree ~ gender, data=dset1, ylab="Degree", xlab="Gender", names=c("Male", "Female"))
dev.off()
## robust t-test
library(WRS2)
t1way(degree ~ gender, data=dset1)

## gender differences in betweenness
t.test(betweenness ~ gender, data=dset1)
pdf("between-boxplot.pdf", width=7, height=7)
boxplot(betweenness ~ gender, data=dset1, ylab="Betweenness", xlab="Gender", names=c("Male", "Female"))
dev.off()
## robust t-test
t1way(betweenness ~ gender, data=dset1)

## gender differences in pagerank
t.test(pagerank ~ gender, data=dset1)
pdf("pagerank-boxplot.pdf", width=7, height=7)
boxplot(pagerank ~ gender, data=dset1, ylab="PageRank", xlab="Gender", names=c("Male", "Female"))
dev.off()
## robust t-test
t1way(pagerank ~ gender, data=dset1)
```

### regression and path analysis


```{r}
View(cor(dset1))
```

Matrix plot of the variables

```{r}
library(lattice)

panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...) {
  usr <- par("usr")
  on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- (cor(x, y, use = "complete.obs"))
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste(prefix, txt, sep = "")
  text(0.5, 0.5, txt, cex =  4*(1 + abs(r)) / 2)
}

panel.hist <- function(x, ...) {
  usr <- par("usr")
  on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 1.5) )
  h <- hist(x, plot = FALSE)
  breaks <- h$breaks
  nB <- length(breaks)
  y <- h$counts
  y <- y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, col = "white", ...)
  box()
}

pdf("matrixplot.pdf", width=7, height=7)
pairs(dset1[,c("degree", "gpa", "bmi", "happy", "lonely")],
      upper.panel = panel.cor,
      diag.panel  = panel.hist,
      lower.panel = panel.smooth)
dev.off()
```

### Mediation analysis

```{r}
library(lavaan)

m1 <-'
degree ~ a1*gpa + a2*bmi
happy ~ c1*gpa + +c2*bmi + b1*degree
lonely ~ c3*gpa + +c4*bmi + b2*degree

## mediation effect
med1 := a1*b1
med2 := a1*b2
med3 := a2*b1
med4 := a2*b2
med5 := (a1+a2)*b1
med6 := (a1+a2)*b2
med7 := a1*b1 + c1
med8 := a1*b2 + c3
med9 := a2*b1 + c2
med10 := a2*b2 + c4
'

m1.res <- sem(m1, data=dset1)

summary(m1.res)
```

## Stochastic block model

```{r}
m.index <- index[nonnetwork$gender==0]
f.index <- index[nonnetwork$gender==1]

order.index <- c(m.index, f.index)

wechat.mat.block <- wechat.mat[order.index, order.index]

length(m.index)

sum(wechat.mat.block[1:74, 1:74])/(74*73)

sum(wechat.mat.block[75:165, 75:165])/(94*93)

sum(wechat.mat.block[1:74, 75:165])/(94*74)
```

### Estimate the SBM

```{r}
library(blockmodels)

sbm1 <- BM_bernoulli("SBM_sym", wechat.mat)
sbm1$estimate()

which.max(sbm1$ICL)

## Plot the ICL
pdf("wechat.ICL.pdf", width=7, height=7)
plot(1:15, sbm1$ICL, xlab="Number of blocks", ylab="ICL")
dev.off()

sbm1$memberships[[10]]$plot()
sbm1$memberships[[10]]$Z
table(sbm1$memberships[[10]]$map()$C) 
table(sbm1$memberships[[7]]$map()$C) 
table(sbm1$memberships[[6]]$map()$C)
table(sbm1$memberships[[5]]$map()$C)

table(sbm1$memberships[[7]]$map()$C, sbm1$memberships[[6]]$map()$C)

round(sbm1$model_parameters[[6]]$pi, 3)

sbm1$memberships[[7]]$map()$C ## class membership
```

```{r}
## plot the data with determined blocks
pdf("wechat.sbm.pdf", width=7, height=7)
par(mar=c(0,0,0,1))
set.seed(3)
layout.coord <- gplot(wechat.network, gmode="graph", usearrows=F, vertex.cex=1.5, vertex.col=sbm1$memberships[[6]]$map()$C, edge.col='grey75', mode="fruchtermanreingold")
legend("topleft", legend=1:6, col=1:6, pch=16, cex = 1.5)
dev.off()

```

## Exponential random graph models

### The model with edge only

```{r}
library(ergm)
t1.model1 <- ergm(wechat.network ~ edges ,
   control=control.ergm(seed=40))

summary(t1.model1)

exp(coef(t1.model1))
plogis(coef(t1.model1))

t1.model1 <- ergm(wechat.network ~ edges + kstar(2) ,
                  estimate = "MPLE",
   control=control.ergm(seed=40))

summary(t1.model1)

t1.model1 <- ergm(wechat.network ~ edges + kstar(2) + triangle,
                  estimate = "MPLE")

summary(t1.model1)

t1.model1 <- ergm(wechat.network ~ edges + kstar(2) + triangle)

summary(t1.model1)


```

### model with node covariates

```{r}
t1.model1 <- ergm(wechat.network ~ edges +
                    nodecov('age') +
                    absdiff('bmi') +
                    nodematch('gender') +
                    nodematch('smoke'),
   control=control.ergm(seed=40))

summary(t1.model1)

model2 <- ergm(wechat.network ~ edges +
                    nodecov('gpa') +
                    absdiff('bmi') +
                    nodematch('gender') +
                    nodematch('smoke', diff=TRUE),
   control=control.ergm(seed=40))

summary(model2)
```

### Latent space models

```{r}
library(latentnet)

## model with the latent position only
latentnet.t1 <- ergmm(wechat.network ~ euclidean(d=1), verbose=TRUE)
mcmc.diagnostics(latentnet.t1)
plot(latentnet.t1, pie=TRUE)

t1.lsm.fit <- summary(latentnet.t1)
t1.lsm.fit

## BIC for model fit

all.fit <- list()
all.fit[[1]] <- t1.lsm.fit$bic

for (d in 2:9){
  latentnet.t1 <- ergmm(wechat.network ~ euclidean(d=d), verbose=TRUE)
  t1.lsm.fit <- summary(latentnet.t1)
  all.fit[[d]] <- t1.lsm.fit$bic
}


all.fit.mat <- do.call("cbind", all.fit)
all.fit.mat

all.fit.mat.nocov <- all.fit.mat

bic29 <- all.fit.mat[6, ]
bic29 <- c(8700.61, bic29)

pdf("latent.space.bic.pdf", width = 7, height = 7)
plot(1:9, bic29, ylab="BIC", xlab="Latent Dimension", type="o")
dev.off()

## plot the results
system.time(latentnet.m2 <- ergmm(wechat.network ~ euclidean(d=2), verbose=TRUE))

#t1.lsm.z <- latentnet.t1$mkl$Z


pdf("latent.space.pdf", width = 7, height = 7)
plot(latentnet.m2, main="", sub="", plot.means=F, plot.vars=F, vertex.col='black')
dev.off()
```


```{r}
system.time(latentnet.t1 <- ergmm(wechat.network ~ euclidean(d=2)+
                    nodecov('gpa') +
                    absdiff('bmi') +
                    nodematch('gender') +
                    nodematch('smoke', diff=TRUE), verbose=TRUE))
mcmc.diagnostics(latentnet.t1)
plot(latentnet.t1, pie=TRUE)

t1.lsm.fit <- summary(latentnet.t1)
t1.lsm.fit
```

